---
title: "HW12: Dates"
author: "Amanda Russom"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

### Load Libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

### Part 1: Play date

## Load personal birth_date chunk
```{r}
russom <- "January 27, 1979"
russom1 <- mdy(russom)
russom1
```

### Question 1

##Elizabeth Anderson
```{r}
anderson <- "9 June 1836"
anderson1 <- dmy(anderson)
anderson1
```

##Mary Anning
```{r}
anning<- "1799/05/21"
anning1 <- ymd(anning)
anning1
```

##Alice Ball
```{r}
ball<- "July 24, 1892"
ball1 <- mdy (ball)
ball1
```

## Sylvia Earle
```{r}
earle <- "August, 30 (1935)"
earle1 <- mdy(earle)
earle1
```

## Rosalind Franklin
```{r}
franklin<- "25th of July, 1920"
franklin1 <- dmy(franklin)
franklin1
```

## Ester Lederberg
```{r}
lederberg <- "December 18, 1922"
lederberg1 <- mdy(lederberg)
lederberg1
```

## Barbara McClintock
```{r}
mclintock <- "16061902"
mcclintock1 <- dmy(mclintock)
mcclintock1
```

## MLK
```{r}
mlk_birth <- "1/15/29"
mlk_birth1 <- mdy(mlk_birth)
year(mlk_birth1) <- 1929
mlk_birth1
```

## MLK
```{r}
mlk_dream <- "1963 August 28"
mlk_dream1 <- ymd(mlk_dream)
mlk_dream1

```

### Question 3
### Weekday of birth
```{r}
weekdays(russom1)
```
### question 4
### Weekday of this year's birthday
```{r}
update(russom1, year= 2020, month=1, day=27) %>% 
  weekdays()
```

### question 5
### 2525 birthday
```{r}
update(russom1, year= 2525, month=1, day=27) %>% 
  weekdays(abbr= TRUE)
```



### set 2525 birthday
```{r}
birthday <- "january 27, 2525"
birthday2525 <- mdy(birthday)
```

### how old in 2525
```{r}
(birthday2525-russom1)/365
```

### the return code provides an answer in days, but divding by 365 converts answer to years

### question6

### days days between Dr. Anderson and Alice Ball
```{r}
ball1-anderson1
```
### days between MLK birth and speech
```{r}
mlk_dream1-mlk_birth1
```

### create birthdates vector
```{r}
birthdates <- c(anderson1, anning1, ball1, earle1, franklin1, lederberg1, mcclintock1, mlk_birth1, russom1)
```
### check for leap years
```{r}
birthdates[leap_year(birthdates)] %>% 
 year()
```
### days between my birthday and the scientist's birthdays 
```{r}
russom1-birthdates
```

