---
title: HW07: Data Importing and Wrangling 1
author: Amanda Russom
output: html_notebook
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

```{r}
file_path <- "../data" 
```

### Load limpets data 

```{r}
limpets_data <- read_csv(file.path(file_path, "limpets.csv"))
```
The limpets data is tidy

#### Use the column types argument


```{r}
col_types = "ncc"
```
### Make the sites column an ordered factor

```{r}
limpets_data$site <- factor(limpets_data$Site, levels=c("PBL", "KNRM", "VBG","WP","PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
```

### Checking order

```{r}
limpets_data
```


### Make a boxplot 
```{r}
ggplot(data = limpets_data, mapping = aes(x = Site, y = Length)) + 
  geom_boxplot(aes(color=Protected)) + labs(x="Site", y="Length (mm)")
```
DP and VBG sites have outliers.  The protected sites are home to larger limpets.


### Roseate Terns

The data is NOT tidy.

### Import tern data 

```{r}
tern_data <- read_tsv(file.path(file_path, "roseate_terns.txt"))
```

### Remove sites with missing data

```{r}
complete_terns <- tern_data %>% filter(complete.cases(.))
```

### Tidy data

```{r}
tern_2 <- complete_terns %>% 
  pivot_longer(AR:JC, names_to = "Sites", values_to = "Populations")  
```

```{r}
tern_2
```


### Tern data graph 

```{r}
tern_2 %>%
  ggplot() +
  geom_line(aes(x = year, y = Populations, group= Sites, color= Sites))
```

Several sites increased in population, including, AR, AL, AC, BL, BC, JL, and JC.  The populations that decreased were BR and JR.

The breaks in the lines are due to missing data which was removed prior to writing the graph code. 


### Import abalone data, remove the first column ans make a boxplot

```{r}
abalone_data <- read_csv(file.path(file_path, "abalone.csv"))
select(abalone_data, -(X1))
ggplot(data  = abalone_data) + 
  geom_boxplot(mapping = aes(x = Type, y = Height))
```

### Make a scatterplot to analyze outliers

```{r}
ggplot(data = abalone_data) + 
  geom_point(mapping = aes(x = Rings, y = Height))
```
The large abalone are not old, according to the ring count data.  Both are </= 10. 

### Filter the data to remove outliers.  Create and ordered factor.  Graph new scatterplot

```{r}
abalone_data_2 <- filter(abalone_data, Height <= 0.35)
factor(abalone_data_2$Type, levels = c("I", "F", "M"))
ggplot(data = abalone_data_2) + 
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```

The new scatter plot shows that mature individuals generally have greater height than immature individuals. However, not all mature male and female individuals are larger than immatures. 

### Create a plot with zero height individuals removed

```{r}
abalone_data_2[abalone_data_2$Height == 0, 'Height'] <- NA
ggplot(data = abalone_data_2) + 
  geom_point(mapping = aes(x = Height, y = Rings, color = Type))
```

 

### Scatterplot of choice #1

```{r}
ggplot(data = abalone_data_2) + 
  geom_point(mapping = aes(x = Rings, y = Diameter, color = Type))
```
### Scatterplot of choice #2

```{r}
ggplot(data = abalone_data_2) + 
  geom_point(mapping = aes(x = Rings, y = LongestShell, color = Type))
```
### Import darters data

```{r}
darter_data <- read_fwf(file.path(file_path, "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = "#")
```

### Name columns, make ordered factors  

```{r}
colnames(darter_data) <- c("Specimen", "Species", "Riffle", "Age", "Sex", "Length", "Depth", "Velocity", "Temp", "MajSub Proportion", "MinSub Proportion", "Major Type", "Minor Type")
darter_data$Riffle <- factor(darter_data$Riffle,levels=c("1", "2"))
darter_data$"Major Type" <- factor(darter_data$"Major Type",levels=c("s","fg", "sg", "lg", "c"), order = TRUE)
```
### Filter rows, Remove and rename columns, arrange data

```{r}
darter_data_2 <- darter_data %>% filter(Species != "zonale" & Species != "tetrazonum") %>% 
  select(!"MinSub Proportion" & ! "Minor Type") %>% 
  rename("Major_substrate" = "MajSub Proportion", "Major_type" = "Major Type") %>% 
  arrange("Specimen")
```
### Darter Plot 1

```{r}
ggplot(data = darter_data_2,mapping = aes(x = Length, y = Depth, color = Species, shape = Species)) +
  geom_point() +
  facet_wrap(~Riffle, ncol = 2) 
```
In riffle 1 the darters lie deeper than in riffle 2.


### Darter Plot 2

```{r}
ggplot(data = darter_data_2,mapping = aes(x = Length, y = Species, color = Species)) +
  geom_boxplot() +
  facet_wrap(~Riffle, ncol = 2)
```
Riffle 2 shows the most outliers.

### Darter Plot #3

```{r}
ggplot(data = darter_data_2, mapping = aes(x = Length, y = Major_type, color = Species))+
  geom_boxplot() +
  facet_grid(Species ~ Riffle)
```
### Switch order of argument

```{r}
ggplot(data = darter_data_2, mapping = aes(x = Length, y = Major_type, color = Species))+
  geom_boxplot() +
  facet_grid(Riffle ~ Species)
```

The argument order of the first graph (species~riffle) allows the data to be less compressed and interpreted more easily