---
title: "HW 09 Wrangling II"
author: "Amanda Russom"
output: html_notebook
editor_options:
  chunk_output_type:iinline
---

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(knitr)
```
```{r}
library(smatr)
```

### Source my_functions
```{r}
source(here("scripts","my_functions.R"))
```

### Import darters.csv and remove species tetrazonum and zonale

```{r}
darters_data <- read_csv(here("Data","darters.csv"))
darters_data_filtered <- darters_data %>% 
  filter(species=="blennioides"|species=="caeruleum"|species=="spectabile")
```
### Sum majsub and minsub to form new column.  Change riffle values.  Change length units. Change sex values.
```{r}
wrangled_darters_data <- mutate(darters_data_filtered, total_substrate=majsub+minsub) %>% 
  mutate(riffle=ifelse(riffle==1,
                       "Riffle 1",
                       "Riffle 2")) %>%
  mutate(length=length*10) %>%
  mutate(sex = ifelse(sex == "f",
                    "Female",
                    "Male"))
  
  
```

### Summarize the mean length, depth, and velocity.  Group by species and riffle.
```{r}
spec_riffle_sep_wrang_darters_data <- group_by(wrangled_darters_data, species, riffle)
summary_wrangled_data <- summarize(spec_riffle_sep_wrang_darters_data, length=mean(length, na.rm=TRUE), depth=mean(depth, na.rm = TRUE),velocity=mean(velocity,na.rm = TRUE))
  
```

### Use kable to make a table
```{r}
kable(summary_wrangled_data)
```
#### Plot 1 summarized darters data
```{r}
ggplot(data = wrangled_darters_data) + 
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```
#### Add geom_point() to add mean length

```{r}
 ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), shape=15)
stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```
### Rotate axis, seperate in to 2 riffle, add labs()

```{r}
ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), shape=15)+
  facet_wrap(~ riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
  
```
# riffle 2 contains the longest average blennioides, but among the other species, the individuals in riffle one are longer


### darters plot 2
### added grayscale and 1st stat_summary
```{r}
ggplot(data = summary_wrangled_data) +
  geom_point(mapping = aes(x=species, y=length), color= "gray65")+
  facet_wrap(~ riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
stat_summary(
    fun.y = median,
    geom = "point",
    size = 2.2
  )
```

### darters plot2, added second stat_summary function to graph

```{r}
ggplot(data = summary_wrangled_data, aes(x= species, y= length))+
  geom_point(color= "gray65")+
stat_summary(
    fun.y = median,
    geom = "point",
    size = 2.2) + 
stat_summary(
    fun.y = mean,
    geom = "point",
    size = 3,
    color = "maroon")+
  facet_wrap(~ riffle) +
  coord_flip() +
  labs(y = "length", x = NULL)
  
  
  
```
```{r}
library(tidyverse)
library(smatr)
```

### load feathers data and remove decimal commas 
```{r}
feathers_data <- read_csv(here("Data","feathers.csv"),locale = locale(decimal_mark = ",") )
```
### Renaming feather columns
```{r}
feathers_data_2 <- feathers_data %>%
  rename("capture_date" = "Capture Date", "day" = "Date","site"= "Site","urban_score"=  "Urba Score","cortF"="CortF ng/mg", "cortB"="CortF ng/mm", "sex"="Sex", "smi"="SMI", "tarsus"="Tarsus", "mass"="Mass", "basal_cort" = "Bas_Cort", "stress_cort" ="Stress_Cort")
```

### calculate the mean tarsus length, store in mean_tarsus

```{r}
mean_tarsus <- mean(feathers_data_2$tarsus, na.rm = TRUE) 
```

### Use mutate () to form new columns 
 
```{r}
mutated_feathers_data_2 <- mutate(feathers_data_2,log_mass=log(mass),log_tarsus=log(tarsus))
```
 
### use sma() function to calculate the standardized major axis regression

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = mutated_feathers_data_2)
```
### apply summary function
```{r}
summary(major_axis)
```
### apply coef() function
```{r}
coef(major_axis)
```

### store slope value in b

```{r}
slope <- coef(major_axis)
b <- slope['slope']
```

### use mutate to add new SMI column
```{r}
feathers_data_2_w_SMI <- mutate(feathers_data_2,new_smi = scaled_mass(mass,tarsus,b)) 
```

#### compare new columns

```{r}
select(feathers_data_2_w_SMI, "smi", "new_smi")
```

### Group by site
```{r}
site_sorted_feathers_data <- group_by(feathers_data_2_w_SMI,site)
```

### calculate the mean and standard error of the mean
```{r}
summarized_feathers_data <- summarise(site_sorted_feathers_data, basal_mean=mean(basal_cort, na.rm = TRUE),stress_mean = mean(stress_cort, na.rm=TRUE), std_err_basal=std_err(basal_cort, na.rm=TRUE),std_err_cort=std_err(stress_cort, na.rm=TRUE))

```
### print results

```{r}
summarized_feathers_data
```
### Plot 1 of feathers data
```{r}
ggplot(data=site_sorted_feathers_data)+
  geom_boxplot(mapping=aes(x=site,y=new_smi))
```


### The individuals of the niort site have much wider range of smi vaules as compared to the other sites

### Plot 2 feathers data
```{r}
ggplot(data=summarized_feathers_data)+
  geom_point(mapping=aes(x=site,y=basal_mean))+
  geom_errorbar(aes(x=site, ymin=std_err_basal-1,ymax=std_err_basal+1))
```

### the assignment wanted us to plot the basal cortico concetration for each site, I hope this only meant the mean vaule and not all the collected vaules, because that's what we did :/










