---
title: "HW10: Data Visualization I"
author: "Amanda Russom"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

### Install and load libraries

```{r}
library(tidyverse)
```

```{r}
library(here)
```

```{r}
library(GGally)
```

```{r}
library(patchwork)
```

### Import Sparrows Data

```{r}
sparrows_raw <- read_tsv(here("data", "sparrows.txt"))
```

## Check
```{r}
sparrows_raw
```

### Wrangle Sparrow Data

```{r}
sparrows <- sparrows_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
```

## Check
```{r}
view(sparrows)
```

### Figure 2: Outliers

## Run Code
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group, y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL, y = "Wing length (mm)")
```
## Run Cleveland dot plot

```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x= wingcrd,y=row_order))+
  geom_point()+
  labs(x="Wing length (mm)", y="Order of the date")
```

## Display plots
```{r}
plot_box + plot_cleveland
```
### Adding coord_flip()
```{r}
plot_box + plot_cleveland+coord_flip()
```
### Figure 3: Outliers

### culmen plot 
```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
 
### nalospi ploy
```{r}
p_nalospi <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
### wt plot
```{r}
p_wt <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
### wingcrd plot
```{r}
p_wingcrd <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
### tarsus plot
```{r}
p_tarsus <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
### head plot
```{r}
p_head <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```
### six variable plot

```{r}
p_culmen+p_nalospi+p_wt+p_wingcrd+p_tarsus+p_head
```

### the variables nalospi, tarsus, and head appear to have outlier data

### figure 5 distribution
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```

# faceted histogram
```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)
```
### patchwork
```{r}
h1+h2
```
### Recreate figure using density
### Create plot 1
```{r}
dens1 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_density(aes(x = wt)) +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```
### Create plot2
```{r}
dens2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_density(aes(x = wt, fill= Month), alpha= 0.50) +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
```

### patching density plots together
```{r}
dens1+dens2
```
### scatterplot panels

```{r}
sparrows %>% ggscatmat(columns = 1:7)
```

### figure 11 plots
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```


