---
title: "HW13: Strings and things"
author: "Amanda Russom"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(here)
library(ggthemes)
```

### Import beta_gliadin data
```{r}
beta_gliadin_raw <- read_lines(here("Data", "beta_gliadin.txt"),skip = 1)
beta_gliadin_raw
```

### replace whitespace and digits with nothing
```{r}
beta_gliadin_cleaned <- str_replace_all(beta_gliadin_raw,"\\d","") %>% 
  str_replace_all("\\s", "")
beta_gliadin_cleaned
  
```
### Collapse the strings
```{r}
beta_gliadin_collapsed <- str_c(beta_gliadin_cleaned, collapse = "")
beta_gliadin_collapsed
```

### Convert to upper case and return length
```{r}
beta_g_upper <- str_to_upper(beta_gliadin_collapsed) 
beta_g_upper
```
### find length of sequence
```{r}
str_length(beta_g_upper)
```


### count the letters that are not A, C, G, T
```{r}
str_count(beta_g_upper, ".A.C.G.T.")
```
### Count start codons 
```{r}
str_count(beta_g_upper, "ATG")
```
### Count stop codons
```{r}
str_count(beta_g_upper, "TGA|TAA")
```
### create caat_box and tata_box variables
```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"

```
### count how many times caat_box occurs in the DNA sequence
```{r}
str_count(beta_g_upper, caat_box)
```
### count how many times tata_box occurs in the DNA sequence
```{r}
str_count(beta_g_upper, tata_box)
```
### create caat_seq search string

```{r}
caat_seq <- "CCAAT.{100,150}ATG"
```
### Count how many times this sequence occurs
```{r}
str_count(beta_g_upper, caat_seq)
```

### Create tata_seq search string
```{r}
tata_seq <- "CTATAA.{50,100}ATG"
```
### Count how many times this sequence occurs
```{r}
str_count(beta_g_upper, tata_seq)
```
### extract all caat_seq, determine the length of each sequence
```{r}
caat_seq_all <- str_extract_all(beta_g_upper, caat_seq, simplify = TRUE)
str_length(caat_seq_all)
```
### use str_which() 
```{r}
str_which(caat_seq_all, tata_seq)

```
### str_which() would only show the postion of the match, not the match itself

### use str_subset() to print the entire match
```{r}
promotor <- str_subset(caat_seq_all, tata_seq)
promotor
```
### determine the length of the match
```{r}
str_length(promotor)
```
### Part 2: Ants

### Import ants data
```{r}
ants_raw <- read_csv(here("Data", "ants.csv"))
ants_raw
```
```{r}
ants_data_wrangled <- ants_raw %>% 
  pivot_longer(names_to="species",
               values_to="abundance",
               matches("[A-Z].+\\..+", ignore.case=FALSE)) %>% 
  select("year", "species", "abundance")
ants_data_wrangled
```
### Rename species
```{r}
ants_mutated <- ants_data_wrangled %>% mutate(species=str_replace(species,"(^[A-Z][a-z]{0,3}).*\\.([a-z]{1,4}).*", "\\1\\2"))
ants_mutated

```
### Create 2010 and 2011 levels, group by year and species and use sum() to add up the abundance of each species 

```{r}
final_ants_data <- ants_mutated %>% 
  mutate(year = factor(year,ordered = TRUE,levels = c(2010, 2011))) %>%
   group_by(year, species) %>%
   summarise(total = sum(abundance))
final_ants_data

```

### plot the data 
```{r}
ggplot(final_ants_data,aes(x = reorder(species,total),y = total),group = species) +
  geom_line() +
  geom_point(aes(shape = year),size = 2) +
  coord_flip() +
  labs(x = NULL, y = "Total Abundance", shape = "Year") +
theme_minimal() 
```
### part 3
### Import and wrangle the first data set
```{r}
flower_size_raw <- read_csv(here("Data", "flower_size.csv")) 
flower_size_raw
```
### filter out flower number na values
```{r}
flower_data_filtered <- drop_na(flower_size_raw, "Flower number")
flower_data_filtered
```
### select and rename columns 
```{r}
wrangled_flower_data <- select(flower_data_filtered, "Species", "Corolla length (cm)", "Corolla width throat (cm)") %>% 
  rename("species" = "Species", "cor_length" = "Corolla length (cm)", "throat_width" = "Corolla width throat (cm)")
wrangled_flower_data

```
### use seperate to seperate genus and species
```{r}
separated_flowers <- wrangled_flower_data %>% 
  separate(species, into = c("genus", "species"))
separated_flowers
```
### group data and use summarize to calculate the mean and max lengths and widths
```{r}
summarized_flower_data <- separated_flowers %>%
  group_by(genus, species) %>% 
  summarize("mean_cor_length" = mean(cor_length, na.rm = TRUE), "mean_throat_width" = mean(throat_width, na.rm = TRUE), "max_cor_length" = max(cor_length, na.rm = TRUE), "max_throat_width"= max(throat_width, na.rm = TRUE))
summarized_flower_data
```
### import and wrangle the second data set

### import data 
```{r}
pollinators_raw <- read_csv(here("Data", "flower_pollinators.csv"))
pollinators_raw
```

### remove columns, remove rows, seperate into genus and species 
```{r}
pollinators <- pollinators_raw %>%
   select(-c(starts_with("Source"), "Color")) %>%
   filter(!is.na(Pollinator)) %>%
   separate(Species, into = c("genus", "species")) 
pollinators
   
```
### seperate pollinator column 
```{r}
syndrome <- pollinators %>% 
 separate(Pollinator, into = "syndrome", sep = "(,| )")
 syndrome

```

### use inner_join
```{r}
phlox_data <- inner_join(summarized_flower_data, syndrome)
phlox_data
```
### phlox plot 1 with improved legend
```{r}
pholox_plot_1 <- phlox_data %>% 
   mutate(syndrome = str_replace(syndrome, "(^be.*e$)", "\\1s")) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = mean_cor_length, y = mean_throat_width, color = syndrome))
pholox_plot_1
  
```

### check syndrome counts
```{r}
phlox_data %>%
  group_by(syndrome) %>% 
  count(syndrome)
```
### phlox plot 2, create other insects in legend
```{r}
phlox_data_2 <- phlox_data %>%
  mutate(syndrome = str_replace(syndrome, "(^be.*e$)", "\\1s"),
         syndrome = str_replace(syndrome, ".+l.?es", "other insects")) 
  
phlox_data_2 %>% 
  ggplot() +
  geom_point(mapping = aes(x=mean_cor_length, y = mean_throat_width, color = syndrome))
  
```

### mutating data

```{r}
mutated_phlox<- phlox_data_2%>% 
mutate(syndrome_group = case_when(
           syndrome == "autogamous" ~ "Autogamous",
           syndrome == "bat" ~ "Mammal",
           syndrome == "hummingbird" ~ "Bird",
           TRUE ~ "Insect"))
mutated_phlox
```

### now plotting data, phlox plot #3
```{r}
mutated_phlox %>% 
  ggplot() +
  geom_point(mapping = aes(x=mean_cor_length, y = mean_throat_width, color = syndrome_group))
```
### improving the plot
```{r}
mutated_phlox %>% 
  ggplot() +
  geom_point(aes(x = max_cor_length,
                 y = max_throat_width,
                 color = syndrome_group)) +
  scale_color_viridis_d(type= "qual", palette = "inferno") +
  labs(x = "Mean Length (cm)",
       y = "Mean Width (cm)",
       color = "Syndrome")
```
### copied from the assignment, but got an error code with scale_color_viridis_d

### 4 plots of my own

## 1
```{r}
mutated_phlox %>%
  ggplot(aes(x=max_cor_length,y=max_throat_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="syndrome",fill="syndrome")+
  theme_tufte ()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_viridis_d(option = "viridis")
```
## 2
```{r}
mutated_phlox %>%
  ggplot(aes(x=max_cor_length,y=max_throat_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)")+
  scale_shape_manual(values=c(20:30))+
  scale_fill_viridis_d(option = "plasma")
```
## 3 
```{r}
mutated_phlox %>% 
  ggplot(aes(x=max_cor_length,y=max_throat_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  theme_economist()+
  scale_shape_manual(values=c(10:45))+
  scale_fill_brewer(palette="Dark1")
```
### 4 
```{r}
mutated_phlox %>% 
  ggplot(aes(x=max_cor_length,y=max_throat_width,shape=syndrome_group,fill=syndrome_group))+
  geom_point()+
  labs(x="Max Length (cm)",y="Max Width (cm)",shape="Syndrome",fill="Syndrome")+
  scale_shape_manual(values=c(10:45))+
  scale_fill_brewer(palette="Dark5")
```
### today was my second day to work on this assignment.  today i worked from 11:00 am until 8:50 pm, it is incredibly overwhelming and defintely not "easy"
